---
title: "datamesh_example_3"
author: "Victor"
date: "2026-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries

```{r}
library(devtools)
library(tidyverse)
```

## Load in Datamesh

```{r}
devtools::load_all()
```

## Connect to Datamesh API
Please keep your Datamesh token hidden.
```{r}
token = Sys.getenv("DATAMESH_TOKEN")
conn <- Connector$new()
```

## Browse catalog

```{r}
catalog <- conn$get_catalog("Whanganui")
catalog
```
## Get details about datasource

```{r}
datasource = conn$get_datasource("flowrateconz_whanganui_river")
datasource
```
## Query data
This is an example of what happens when we timefilter outside the applicable time range, the filter will return NA on requested times with no data.
```{r}
query_input = Query$new(datasource = "flowrateconz_whanganui_river",
                  timefilter = list(times = c("2024-01-1", "2025-01-1")))
data = conn$query(query_input)
              
```

## Format the data
The time format is returned as a string which we need to convert into a plot-able time format.
```{r}
flow = data$data_vars$flow$data
rain = data$data_vars$rain$data
time = as.POSIXct(data$coords$time$data, format = "%Y-%m-%dT%H:%M:%S", tz = "UTC")
```

```{r}
df <- tibble(time = time, flow = flow, rain = rain)
```

## Plot data

```{r}
df |> ggplot(aes(x = time, y = flow)) +
  geom_line() + 
  labs(x = "Time", 
       y = bquote("Average Hourly Flow Rate (" ~ m^3/s ~ ")"),
       title = "Whanganui River Average Hourly Flow Rate between 2024 and 2025")
```


```{r}
df |> ggplot(aes(x = time, y = rain)) +
  geom_line() + 
  labs(x = "Time", 
       y = bquote("Average Hourly Rain (" ~ mm ~ ")"),
       title = "Whanganui River Average Hourly Rain between 2024 and 2025")
```

