---
title: "datamesh_example_1"
author: "Victor"
date: "2026-01-22"
output: html_document
---

## Import Relevant Libraries

```{r}
library(devtools)
library(tidyverse)
```

## load Datamesh

```{r}
devtools::load_all()
```
## Connect to Datamesh API
By default Connector will grab the system variable DATAMESH_TOKEN however a datamesh token can also be passed in.

```{r}
token <- Sys.getenv("DATAMESH_TOKEN")
conn <- Connector$new(token = token)
```

## Browse catalog
There are a lot of datasources in the "oceanum" catalog so we can limit the amount found using the limit.
```{r}
catalog <- conn$get_catalog(search = "oceanum", limit = 15)
catalog
```

## Getting metadata about datasource
Once we've found a datasource we can pull more information about the datasource using get_datasource.

```{r}
datasource <- conn$get_datasource("oceanum-sea-level-rise")
datasource

```

## Query datasource to load it into work environment.

```{r}
query <- Query$new(datasource = "oceanum-sea-level-rise")
data <- conn$query(query_input = query)
```

## Clean data 
Once the data is loaded in, the formatting needs to be cleaned so that the data is tidy.
```{r}
clean_data <- tibble(days = as.Date(data$data_vars$Day$data), 
               average = data$data_vars$sea_level_rise_average$data)
```

## Plot data
R has inbuilt plotting function however personally I prefer GGplot2 for aesthetics and customisation. 
```{r}
clean_data |> ggplot(aes(x = days, y = average)) + 
  geom_point() + 
  theme_gray() + 
  labs(title = "Oceanum Sea Level Rise Example",
       y = "Average sea level rise",
       x = "Year")
```


