---
title: "datamesh_example_2"
author: "Victor"
date: "2026-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Relevant Libraries

```{r}
library(devtools)
library(tidyverse)
library(viridis)
library(maps)
```

## load Datamesh

```{r}
devtools::load_all()
```

## Connect to Datamesh API
By default Connector will grab the system variable DATAMESH_TOKEN however a datamesh token can also be passed in. 
You can set your Datamesh token using: sys.setenv(DATAMESH_TOKEN = "your_token")
```{r}
conn = Connector$new()
```

## Browse catalog
```{r}
catalog = conn$get_catalog(search = "ERA5 wind", limit = 20)
catalog
```

## Get datasource information
```{r}
datasource = conn$get_datasource("era5_wind10m")
datasource
```
```{r}
names(datasource$variables)
```
## Query the data into our work environment
query also works for list inputs
```{r}
query = list(datasource = "era5_wind10m",
             timefilter = list(times = c("2020-01-01", "2020-01-01")))
data = conn$query(query)

```

## Working with the data
Save data into dataframe for plotting
```{r}
u10 <- data$data_vars$u10$data
v10 <- data$data_vars$v10$data

lon <- data$coords$longitude$data
lat <- data$coords$latitude$data


speed <- sqrt(u10^2 + v10^2)
speed2d <- speed[1, , ]

stopifnot(
  nrow(speed2d) == length(lat),
  ncol(speed2d) == length(lon)
)

lon_fixed <- lon
lon_fixed[lon_fixed > 180] <- lon_fixed[lon_fixed > 180] - 360

lon_ord <- order(lon_fixed)
lon_fixed <- lon_fixed[lon_ord]
speed2d <- speed2d[ , lon_ord]

lat_ord <- order(lat)
lat_fixed <- lat[lat_ord]
speed2d <- speed2d[lat_ord, ]

df <- expand_grid(
  lon = lon_fixed,
  lat = lat_fixed
) %>%
  mutate(
    speed = as.vector(speed2d)
  )
c(min(speed),max(speed))
```

## Visualising the data
Let plot the wind speed using ggplot
``` {r}
ggplot(df, aes(lon, lat, fill = speed)) +
  geom_raster() +
  #borders("world", colour = "black", fill = NA) +
  coord_fixed(expand = FALSE) +
  scale_fill_viridis_c(
    name = "Wind speed (m s⁻¹)",
    limits = c(0, 25),
    option = "turbo"
  ) +
  labs(
    title = "ERA5 10 m Wind Speed",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_minimal()

```


